### 2019-06-03 Notes/SRG_operator_evolution/srg_operator_evolution.tex
Added SRG-evolved potential figures to these SRG operator evolution notes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% File: srg_operator_evolution.tex
%
% Author: A. J. Tropiano (tropiano.4@osu.edu)
% Date: May 20, 2019
%
% Notes on SRG operator evolution project.
%
% Revision history:
% 	05/31/19 --- Renamed magnus.tex from magnus_Jan08.tex. - Renamed to srg_operator_evolution.tex from operator_evolution_May20.tex.
%			   Added a description of computing SRG unitary transformations. Added momentum projection operator contours and momentum
%                         distribution plots.
%	06/03/19 --- Added brief descriptions of how to numerically compute momentum projection operators and distributions. Also added SRG-
%        		   evolved potential contours for comparison with other operators.                  
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[preprintnumbers,floatfix,aps,prc,preprint,nofootinbib]{revtex4-1}

% Packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}
\usepackage[font=small,skip=0pt]{caption} % For captions on figures and tables
\usepackage{cellspace}
\usepackage{color}
\usepackage{enumerate}
\usepackage{epsfig}
\usepackage[figuresright]{rotating}
\usepackage{float}
\usepackage{hyperref} % For clickable links to sections within table of contents
\usepackage{graphicx}
\graphicspath{{Figures/}} % Setting the graphics path
\usepackage{physics} % For bra-ket notation
\usepackage{siunitx}
\usepackage[caption=false]{subfig} % For sub-figures

\newcommand{\eps}{\varepsilon}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{SRG operator evolution}

\author{A.~J.~Tropiano$^{1}$, S.~K.~Bogner$^{2}$, R.~J.~Furnstahl$^{1}$}

\affiliation{
		$^1$\mbox{Department of Physics, The Ohio State University, Columbus, OH 43210, USA}  \\
		$^2$\mbox{National Superconducting Cyclotron Laboratory and Department of Physics and Astronomy,}  \\
		\mbox{Michigan State University, East Lansing, MI 48824, USA}
}

\date{\today}

\begin{abstract}

Brief description of project.

\end{abstract}

\maketitle

\tableofcontents

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}

Results on SRG-evolved operators from several NN potentials.
\\
-- How operators evolve from band- and block-diagonal SRG transformations.
\\
-- Operator evolution for different potentials (regulators, chiral order, etc.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mathematical/computational details}
\label{sec:math_comp_details}

%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Building SRG unitary transformations}
\label{sec:srg_unitary_transformations}

Diagonalize initial and evolved Hamiltonians which we will call $H(0)$ and $H(s)$, respectively. This gives $\psi_{\alpha}(0)$ and $\psi_{\alpha}(s)$ for each eigenvalue indexed by $\alpha$. Then the SRG unitary transformation can be computed by taking a sum over outer products of the evolved and initial wave functions:

\begin{eqnarray}
	\label{eq:unitary_transformation}
	U(s) = \sum_{\alpha=1}^{N} \ket{\psi_{\alpha}(s)} \bra{\psi_{\alpha}(0)},
\end{eqnarray}

where N is the dimension of the Hamiltonian matrix. Here the weights are factored into the wave functions, thus $U(s)$ is unitless.
\\

To evolve operators, we simply apply $U(s)$:

\begin{eqnarray}
	\label{eq:evolved_operator}
	O(s) = U(s) O(0) U^{\dagger}(s),
\end{eqnarray}

where $O(0)$ is the bare operator.

%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Momentum projection operator: $a^{\dagger}_q a_q$}
\label{sec:momentum_proj_operator}

Applying $a^{\dagger}_q a_q$ to a wave function $\psi(k)$ returns $\psi(q)$. For the discrete case, $\psi(k_i)$ is an $N \times 1$ vector and $a^{\dagger}_q a_q (k_i, k_j)$ is an $N \times N$ matrix where $i$, $j=1$, ... , $N$. Then $a^{\dagger}_q a_q$ acting on $\psi(k)$ is a matrix multiplication, implying a continuous integration over $d^3k$. Therefore, we include a factor of $1 / (k^2 dk) \implies 1 / (k_i k_j \sqrt{w_i w_j})$ in $a^{\dagger}_q a_q (k_i, k_j)$ where $w$ represents the momentum weights. In matrix form,

\begin{eqnarray}
	\label{eq:momentum_projection_operator}
	a^{\dagger}_q a_q (k_i, k_j) = \frac{\delta_{k_i q} \delta_{k_j q}}{k_i k_j \sqrt{w_i w_j}},
\end{eqnarray}

which has units fm$^3$. To evolve operators, we apply $U(s)$ at this point. For mesh-independent figures, we must divide by an additional factor of $k_i k_j \sqrt{w_i w_j}$.

%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Momentum distribution function: $\phi^2(k)$}
\label{sec:momentum_dist_funcs}

We diagonalize the Hamiltonian for eigenvectors $\psi_{\alpha}$. In the $^{3}S_1$-$^{3}D_1$ coupled channel, the S-component is given by $\psi_{\alpha}[:N]$ and the D-component by $\psi_{\alpha}[N:]$ where $N$ is the length of the momentum mesh. Then the momentum distribution of the state $\alpha$ is given by,

\begin{eqnarray}
	\label{eq:momentum_distribution}
	|\phi_{\alpha}(k)|^2 = |\psi_{\alpha}[:N]|^2 + |\psi_{\alpha}[N:]|^2.
\end{eqnarray}

This satisfies the normalization condition $\sum_{i=1}^N |\phi(k_i)|^2 = 1$, implying that the factor $k^2 dk$ (or in the discrete case, $k_i^2 w_i$) is factored into the wave function. For mesh-independent figures, divide by $k_i^2 w_i$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

Organize this according to the figures. Format should be description of the calculation (previous section), followed by the figure, followed by takeaways.

%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Entem-Machleidt N$^3$LO non-local potential}
\label{sec:nonlocal_results}

Non-local potential from \cite{Entem:2003ft}.

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn10_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn10_3S1_Block-diag2,00}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn10_3S1_Block-diag3,00}%
	}
	\caption{Caption.}
	\label{potential_contours_kvnn10_3S1}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn10_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn10_3S1_Block-diag2,00}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn10_3S1_Block-diag3,00}%
	}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the Entem-Machleidt N$^3$LO non-local potential with the Wegner generator (a) and block-diagonal generators decoupling at $\Lambda=2$ and $3$ fm$^{-1}$ (b and c). Here $q=3$ fm$^{-1}$.}
	\label{momentum_projection_contours_q3,00_kvnn10}
\end{figure}

-- The top row of Fig.~\ref{momentum_projection_contours_q3,00_kvnn10} should match the top row in Fig. 4 of \cite{Anderson:2010aq}.
-- Smeared delta function comment.
-- Compared block-diagonal to Wegner.
-- Another note about block-diagonal.

\begin{figure}[H]
	\centering
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q0,30_kvnn10_3S1_Wegner}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the Entem-Machleidt N$^3$LO non-local potential with the Wegner generator. Here $q=0.3$ fm$^{-1}$.}
	\label{momentum_projection_contours_q0,30_kvnn10}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{deuteron_momentum_distribution_kvnn10_Wegner_lamb1,5}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{deuteron_momentum_distribution_kvnn10_Block-diag2,00_lamb1,5}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{deuteron_momentum_distribution_kvnn10_Block-diag3,00_lamb1,5}%
	}
	\caption{Momentum probability densities of the deuteron SRG-evolving the wave function to $\lambda=1.5$ fm$^{-1}$ from the Entem-Machleidt N$^3$LO non-local potential with the Wegner generator (a) and block-diagonal generators decoupling at $\Lambda=2$ and $3$ fm$^{-1}$ (b and c). The black dotted line corresponds to the momentum probability density of the initial deuteron wave function.}
	\label{deuteron_momentum_distribution_kvnn10}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{continuum_state_momentum_distribution_eps200,0_kvnn10_Wegner_lamb1,5}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{continuum_state_momentum_distribution_eps200,0_kvnn10_Block-diag2,00_lamb1,5}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{continuum_state_momentum_distribution_eps200,0_kvnn10_Block-diag3,00_lamb1,5}%
	}
	\caption{Momentum probability densities of the continuum state at $\epsilon \approx 200$ MeV SRG-evolving the wave function to $\lambda=1.5$ fm$^{-1}$ from the Entem-Machleidt N$^3$LO non-local potential with the Wegner generator (a) and block-diagonal generators decoupling at $\Lambda=2$ and $3$ fm$^{-1}$ (b and c). The black dotted line corresponds to the initial momentum probability density.}
	\label{continuum_state_momentum_distribution_eps200,0_kvnn10}
\end{figure}

%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{RKE N$^3$LO and N$^4$LO semi-local potentials}
\label{sec:semilocal_results}

\textcolor{red}{Add takeaways for these figures.} Semi-local potentials from \cite{Reinert:2017usi}.

% RKE N3LO results

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn106_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn106_3S1_Block-diag2,00}%
	}
	\caption{Caption.}
	\label{potential_contours_kvnn106_3S1}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn107_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn107_3S1_Block-diag2,00}%
	}
	\caption{Caption.}
	\label{potential_contours_kvnn111_3S1}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn106_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn106_3S1_Block-diag2,00}%
	}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the RKE N$^3$LO semi-local potential with the Wegner generator (a) and block-diagonal generator decoupling at $\Lambda=2$ fm$^{-1}$ (b). Here $q=3$ fm$^{-1}$ and the EFT cutoff is $450$ MeV.}
	\label{momentum_projection_contours_q3,00_kvnn106}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn107_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn107_3S1_Block-diag2,00}%
	}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the RKE N$^3$LO semi-local potential with the Wegner generator (a) and block-diagonal generator decoupling at $\Lambda=2$ fm$^{-1}$ (b). Here $q=3$ fm$^{-1}$ and the EFT cutoff is $500$ MeV.}
	\label{momentum_projection_contours_q3,00_kvnn107}
\end{figure}

% RKE N4LO results

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn111_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn111_3S1_Block-diag2,00}%
	}
	\caption{Caption.}
	\label{potential_contours_kvnn111_3S1}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn112_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn112_3S1_Block-diag2,00}%
	}
	\caption{Caption.}
	\label{potential_contours_kvnn112_3S1}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn111_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn111_3S1_Block-diag2,00}%
	}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the RKE N$^4$LO semi-local potential with the Wegner generator (a) and block-diagonal generator decoupling at $\Lambda=2$ fm$^{-1}$ (b). Here $q=3$ fm$^{-1}$ and the EFT cutoff is $450$ MeV.}
	\label{momentum_projection_contours_q3,00_kvnn111}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn112_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn112_3S1_Block-diag2,00}%
	}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the RKE N$^4$LO semi-local potential with the Wegner generator (a) and block-diagonal generator decoupling at $\Lambda=2$ fm$^{-1}$ (b). Here $q=3$ fm$^{-1}$ and the EFT cutoff is $500$ MeV.}
	\label{momentum_projection_contours_q3,00_kvnn112}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{continuum_state_momentum_distribution_eps200,0_kvnn111_Wegner_lamb1,5}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{continuum_state_momentum_distribution_eps200,0_kvnn111_Block-diag2,00_lamb1,5}%
	}
	\quad
	\subfloat[]{%
	\includegraphics[clip,width=0.3\columnwidth]{continuum_state_momentum_distribution_eps200,0_kvnn111_Block-diag3,00_lamb1,5}%
	}
	\caption{Momentum probability densities of the continuum state at $\epsilon \approx 200$ MeV SRG-evolving the wave function to $\lambda=1.5$ fm$^{-1}$ from the RKE N$^4$LO semi-local potential with the Wegner generator (a) and block-diagonal generators decoupling at $\Lambda=2$ and $3$ fm$^{-1}$ (b and c). The black dotted line corresponds to the initial momentum probability density. Here the EFT cutoff is $450$ MeV.}
	\label{continuum_state_momentum_distribution_eps200,0_kvnn111}
\end{figure}

%------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Gezerlis N$^2$LO local potentials}
\label{sec:local_results}

\textcolor{red}{Add takeaways for these figures.} Local potentials from \cite{Gezerlis:2014zia}.

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn222_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn222_3S1_Block-diag2,00}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn222_3S1_Block-diag3,00}%
	}
	\caption{Caption.}
	\label{potential_contours_kvnn222_3S1}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn224_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{potential_contours_kvnn224_3S1_Block-diag2,00}%
	}
	\caption{Caption.}
	\label{potential_contours_kvnn224_3S1}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn222_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn222_3S1_Block-diag2,00}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn222_3S1_Block-diag3,00}%
	}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the Gezerlis et al. N$^2$LO local potential with the Wegner generator (a) and block-diagonal generators decoupling at $\Lambda=2$ and $3$ fm$^{-1}$ (b and c). Here $q=3$ fm$^{-1}$ and the EFT cutoff is $1$ fm.}
	\label{momentum_projection_contours_q3,00_kvnn222}
\end{figure}

\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn224_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{momentum_projection_contours_q3,00_kvnn224_3S1_Block-diag2,00}%
	}
	\caption{Matrix elements of $\mel{k}{a^{\dagger}_q a_q}{k'}$ SRG-evolving in $\lambda$ right to left under transformations from the Gezerlis et al. N$^2$LO local potential with the Wegner generator (a) and block-diagonal generator decoupling at $\Lambda=2$ fm$^{-1}$ (b). Here $q=3$ fm$^{-1}$ and the EFT cutoff is $1.2$ fm.}
	\label{momentum_projection_contours_q3,00_kvnn224}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{../tropiano_bib}

\end{document}